\documentclass{article}

% NeurIPS 2026 — Evaluations & Datasets Track, anonymised review build.
% After acceptance, swap "[eandd]" for "[eandd, final]". For an arXiv preprint
% with author names, use "[eandd, preprint]" — never combine "preprint" with
% "final".
\usepackage[eandd]{neurips_2026}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath,amssymb}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}

\title{Sovereign Regulatory NLP for Indian BFSI:\\A Deterministic Pipeline, Per-Customer LoRA Adapters,\\and Honest Out-of-Distribution Disclosure}

\author{%
  Anonymous Authors \\
  Anonymous Affiliation \\
  \texttt{anonymous@example.com} \\
}

\begin{document}

\maketitle

\begin{abstract}
Sovereign deployment of regulatory AI in Indian banking, financial services, and insurance (BFSI) faces two structural constraints: data-localisation rules restrict cloud APIs, and per-customer compliance corpora differ in style and structure from any single general LLM's training distribution. We contribute (i) a deterministic, contamination-resistant pipeline that scrapes 130 public Reserve Bank of India (RBI) Master Directions and Securities and Exchange Board of India (SEBI) Master Circulars, constructs 4{,}477 question--answer pairs through a 3-tier regex template grammar, and gates every candidate through a 10-check validator (98.45\% pass); (ii) a \emph{document-disjoint} train/eval split that quarantines 26 entire PDFs (20\%) before any pair is constructed, eliminating chunk-neighbour memorisation; and (iii) an open 60-question hand-curated paired benchmark released under CC-BY-SA-4.0 with a gated reference scorer. On 664 held-out paired questions, a single 4-bit LoRA adapter trained in 3h~28min on a single consumer GPU lifts substring match from 58.7\% to 89.6\% (+30.9~pp; McNemar $b_{10}{=}14$, $b_{01}{=}219$, $p{=}1.66\times10^{-48}$). The same recipe transfers to a structurally different no-context recall task ($n{=}214$, $+38.4\%$ relative F1, Wilcoxon $p{=}1.50\times10^{-16}$). On the contemporary public benchmark IndiaFinBench~\citep{pall2026indiafinbench}, the same adapter scores 32.1\% (Wilson 95\% CI [27.3, 37.4]) versus 89.7\% reported for Gemini~2.5 Flash---a 57.6~pp gap we frame as the central evidence \emph{for} per-customer training rather than against the methodology. On the released 60-question benchmark the adapter scores 50\% versus 40\% (McNemar $p{=}0.0313$). Code, data manifests, scorer, and evaluation JSONLs are released.
\end{abstract}

\section{Introduction}

The Indian BFSI sector is governed by a thicket of regulators---RBI, SEBI, IRDAI, PFRDA, and the FEMA framework---that issue thousands of Master Directions, Master Circulars, and amendments each year. Compliance teams at banks, NBFCs, asset managers, and insurance brokers must locate, interpret, and cite specific paragraphs from these documents under audit pressure. Two structural constraints shape any deployable AI assistant in this market.

First, India's Digital Personal Data Protection Act and sectoral RBI circulars on cloud usage place strong pressure toward on-premise or sovereign-cloud deployment. Many regulated entities are unwilling to send customer or transaction data to general-purpose frontier APIs, and many are operationally restricted to on-device or regional inference. This rules out the typical recipe of ``few-shot a frontier model behind a thin retrieval layer'' for a sizeable fraction of the addressable market.

Second, every customer's compliance corpus differs in style. RBI Master Directions are heading-numbered, paragraph-extractive, and verbose; SEBI Master Circulars are tabular and clause-numbered; an individual NBFC's internal compliance handbook adds yet another stylistic layer on top. A single fine-tuned model that wins on one customer's questions can fail on another's---not because the model has lost knowledge, but because the answer surface form has shifted.

We argue these constraints together motivate \emph{per-customer adapters}: small (1--2\% trainable parameters) LoRA adapters trained on each customer's regulatory corpus, served from a single base model on commodity hardware, with rigorous paired evaluation against a base baseline before deployment. The contribution of this paper is the methodology---pipeline, splits, statistical protocol, and released artefacts---not any one trained adapter. We make four concrete contributions.

\begin{enumerate}
  \item \textbf{A deterministic, contamination-resistant pipeline} that turns 130 public regulator PDFs into 4{,}477 raw QA pairs without LLM-generated questions, with provenance preserved at chunk, page, and section granularity (Section~\ref{sec:methods}).
  \item \textbf{A document-disjoint paired evaluation protocol}, applied at $n{=}664$ extract and $n{=}214$ recall, that eliminates chunk-neighbour memorisation. Substring match lifts $+30.9$~pp at $p{=}1.66\times10^{-48}$ (Section~\ref{sec:results}).
  \item \textbf{Honest out-of-distribution disclosure} on IndiaFinBench~\citep{pall2026indiafinbench}: the same adapter scores 32.1\% vs Gemini Flash's reported 89.7\%. We argue this gap supports rather than undermines the per-customer thesis (Sections~\ref{sec:results}, \ref{sec:discussion}).
  \item \textbf{An open 60-question hand-curated benchmark} (CC-BY-SA-4.0) with paired baselines, alternative-answer lists, and a gated reference scorer including built-in Wilson confidence intervals and paired McNemar.
\end{enumerate}

We position the work as an evaluations-and-datasets contribution: methodology, dataset construction, paired statistical protocol, and the released benchmark are central; the OOD result is treated as an empirical claim about distribution shift, not as an apples-to-apples performance comparison.

\section{Background and related work}

\paragraph{Domain adaptation via parameter-efficient fine-tuning.} Low-rank adaptation (LoRA) injects rank-$r$ updates into attention and MLP weight matrices of a frozen backbone, yielding a small trainable module per domain that is typically 1--2\% of the base parameter count~\citep{hu2021lora}. QLoRA extends this idea to 4-bit quantised backbones, enabling fine-tuning of 30B--65B models on a single GPU by combining NF4 quantisation, double quantisation, and paged optimisers~\citep{dettmers2023qlora}. We use these as off-the-shelf primitives; our contribution is not a new fine-tuning method but a faithful, contamination-resistant evaluation of a well-tuned LoRA adapter on a regulator-specific QA task.

\paragraph{Data contamination in QA evaluation.} Several works document contamination of downstream benchmarks in pretraining corpora and the resulting memorisation and exploitation effects~\citep{magar2022data}. Beyond pretraining-level contamination, \citet{magar2022data} and others highlight that even within a fixed corpus, random-row train/test splits can enable ``chunk-neighbour'' memorisation: an answer span memorised in one paragraph supports correct generation in a paraphrased neighbour. Standard random-row splits on regulatory QA inherit this pathology. Our document-disjoint split (Section~\ref{sec:split}) is a direct response.

\paragraph{Out-of-distribution evaluation of fine-tuned models.} Holistic evaluation work such as HELM~\citep{liang2022helm} repeatedly finds that models tuned on one benchmark style transfer poorly to differently-styled questions on the same domain. Our IndiaFinBench result is consistent with this literature; we make the OOD gap the centre of the discussion rather than an appendix.

\paragraph{Regulatory NLP and Indian BFSI benchmarks.} IndiaFinBench is, to our knowledge, the first publicly available benchmark for evaluating LLMs on Indian financial regulatory text~\citep{pall2026indiafinbench}. It comprises 406 expert-annotated QA pairs from 192 SEBI and RBI documents spanning 1992--2026, with four task types: regulatory interpretation, numerical reasoning, contradiction detection, and temporal reasoning. The IndiaFinBench authors evaluate twelve models under zero-shot conditions and find that frontier models such as Gemini 2.5 Flash reach 89.7\% overall accuracy. We use the $n{=}324$ IndiaFinBench test split as our OOD probe. To our knowledge no other open Indian-BFSI benchmark with paired baselines existed at the time of our adapter training; our 60-question benchmark complements IndiaFinBench by providing paired-evaluation baselines under CC-BY-SA-4.0.

\section{Methods and system design}
\label{sec:methods}

\subsection{Corpus}

We scrape 80 RBI Master Directions and 50 SEBI Master Circulars (130 PDFs, 115~MB), all in the public domain. Text is extracted with \texttt{pdfplumber} and \texttt{PyMuPDF} under multiprocessing, yielding 8.06~M characters. We detect 7{,}329 numbered section boundaries and emit 4{,}185 chunks with median 384 tokens via a smart chunker that preserves clause numbering. Provenance for every chunk is logged: source PDF, page span, section heading.

\subsection{Three-tier deterministic QA construction}

We construct QA pairs deterministically from regulator text using three regex template families. We deliberately avoid LLM-generated questions to prevent paraphrase contamination of a downstream LLM evaluator and to preserve full provenance for auditability.

\begin{itemize}
  \item \textbf{Tier 1 (native FAQ, $n{=}1{,}142$).} Native ``Q./A.'' style FAQ blocks already present in some Master Directions are pulled directly with a strict template.
  \item \textbf{Tier 2 (numeric extractive, $n{=}2{,}219$).} Sentences containing numeric facts (rupee amounts, days, percentages, ratios, ISIN/PAN-style identifiers) are converted into ``\emph{What is the X for Y?}''-style questions where the numeric token is the gold answer. Patterns are anchored to surrounding clause structure.
  \item \textbf{Tier 3 (heading-extractive, $n{=}1{,}116$).} Section/sub-section headings are converted into ``\emph{What does \textless heading\textgreater{} say?}'' questions whose gold answer is the corresponding paragraph.
\end{itemize}

This yields 4{,}477 raw QA pairs.

\subsection{Ten-check validator}

Each candidate pair passes a 10-check validator before it is allowed into the dataset. Checks include: question/answer minimum and maximum length, gold answer present in source chunk, chunk--PDF--section provenance fields populated, no markup leakage in answer span, language-detector pass for Hindi/Devanagari spillover, and basic sanity heuristics (no all-caps blocks, no orphan footnote markers). Pass rate after a v2 cleaner pass is 98.45\%. Failed candidates are discarded, not paraphrased; this preserves the determinism of the pipeline.

\subsection{Document-disjoint train/eval split}
\label{sec:split}

We split at the PDF level, not at the QA-pair level: 104 PDFs (80\%) are assigned to train and 26 PDFs (20\%) are quarantined as held-out. Every QA pair is then routed to train/eval based solely on which PDF its context came from. This eliminates the chunk-neighbour memorisation pathway~\citep{magar2022data}: no eval-time paragraph appears (paraphrased or otherwise) in the train set. The final paired dataset is 2{,}931 train + 664 eval pairs, with split manifest stored at \texttt{data/rbi\_corpus/qa/split\_manifest\_v2.json}.

A natural concern is whether a random-row split on the same 4{,}477 raw QA pairs would have inflated apparent gains. We argue (and document in the limitations) that a random-row split would have placed paraphrased neighbours of held-out paragraphs into the train set, allowing the model to memorise span--answer associations rather than learn extractive behaviour. Document-disjoint splitting is the protocol-level fix, not a hyperparameter.

\subsection{BFSI extract adapter}

We train a single LoRA adapter (\texttt{bfsi\_extract}) on top of Nemotron-Nano-30B-A3B (4-bit NF4):

\begin{itemize}
  \item \textbf{Architecture.} LoRA $r{=}16$, $\alpha{=}32$, target modules over both attention and MLP projections. Trainable parameters: 434.6~M of 32~B = 1.36\%.
  \item \textbf{Training.} One epoch, paged AdamW 8-bit, learning rate $2\times10^{-4}$ with cosine schedule, sequence length 1024, on a single RTX~5090.
  \item \textbf{Cost.} Wall-clock 3h~28min, peak VRAM 17.81~GB, adapter file size 1.74~GB.
\end{itemize}

\subsection{BFSI recall adapter}

To probe whether the recipe generalises beyond extractive-with-context QA, we build a second dataset of \emph{no-context recall} questions ($n{=}214$ paired, document-disjoint) where the gold answer is a numeric or short factual span and the question must be answered without an accompanying paragraph. We train an attention-only LoRA (same $r{=}16$, $\alpha{=}32$, sequence length 384, all other settings identical) on the recall train split.

\subsection{IndiaFinBench OOD probe}

For the external OOD probe, we evaluate the same \texttt{bfsi\_extract} adapter---no further fine-tuning, no prompt engineering, identical decoding settings---on the IndiaFinBench $n{=}324$ test split~\citep{pall2026indiafinbench}. Sample contexts in IndiaFinBench range 200--8000 tokens, longer than our internal training contexts (1024). Evaluation runs in 73~min on a single RTX~5090.

\subsection{Indian BFSI Benchmark v1}

We hand-curate a small open benchmark of 60 questions (RBI 30, SEBI 30; Tier 2 numeric 30, Tier 3 heading-extractive 30) drawn from 22 distinct held-out PDFs. Each question carries a hand-curated alternative-answer list (e.g., ``Rs.~1{,}00{,}000'', ``one lakh rupees'', ``INR 100000'') so that scoring is robust to surface variation. The benchmark uses a mixed metric: 30 questions scored by substring match, 30 scored by token-F1 with threshold 0.5. We release the data, scorer, and metadata under CC-BY-SA-4.0; the scorer ships with built-in Wilson confidence intervals and paired McNemar exact-binomial tests.

\subsection{Evaluation metrics and statistical protocol}

Across all evaluations we report case-insensitive substring match as the primary citation-faithful metric, token-F1 as a complementary semantic metric, and Wilson 95\% confidence intervals on the binary substring rate. For paired comparisons we report the McNemar $2\times2$ contingency $(b_{11}, b_{10}, b_{01}, b_{00})$ together with the exact-binomial $p$-value computed via SciPy's \texttt{binomtest}. Where token F1 is the primary outcome (Section~\ref{sec:results-recall}), we use the Wilcoxon signed-rank test for paired F1 differences. All paired evaluations share base decoding settings (temperature 0, identical \texttt{max\_new\_tokens}, identical prompt template) and identical seeds. We do not re-tune for any single benchmark.

\section{Experiments}

We run three paired evaluations and one out-of-distribution probe.

\begin{itemize}
  \item \textbf{E1 -- BFSI extract held-out ($n{=}664$).} Two modes: base, \texttt{bfsi\_extract} direct.
  \item \textbf{E2 -- BFSI recall held-out ($n{=}214$).} Two modes: base, \texttt{bfsi\_recall} direct. Primary metric token F1.
  \item \textbf{E3 -- IndiaFinBench OOD ($n{=}324$).} \texttt{bfsi\_extract} only, evaluated against published Gemini Flash zero-shot scores from IndiaFinBench~\citep{pall2026indiafinbench}.
  \item \textbf{E4 -- Indian BFSI Benchmark v1 ($n{=}60$).} Two modes (base, \texttt{bfsi\_extract}) on the released benchmark.
\end{itemize}

\section{Results}
\label{sec:results}

\subsection{BFSI extract on held-out paired ($n{=}664$)}

Table~\ref{tab:extract-main} reports substring and token-F1 scores.

\begin{table}[h]
  \caption{Held-out paired evaluation of the \texttt{bfsi\_extract} adapter, $n{=}664$ document-disjoint. The adapter lifts substring match from 58.7\% to 89.6\% (+30.9~pp); paired McNemar $b_{10}{=}14$, $b_{01}{=}219$, $p{=}1.66\times10^{-48}$.}
  \label{tab:extract-main}
  \centering
  \begin{tabular}{lccc}
    \toprule
    Mode & Substring & Wilson 95\% CI & Token F1 \\
    \midrule
    Base Nemotron-30B & 58.7\% & [54.95, 62.42] & 0.132 \\
    + \texttt{bfsi\_extract} & \textbf{89.6\%} & [87.06, 91.71] & 0.172 \\
    \bottomrule
  \end{tabular}
\end{table}

The adapter--vs--base McNemar contingency is $(b_{11}, b_{10}, b_{01}, b_{00}) = (376,14,219,55)$, giving $p{=}1.66\times10^{-48}$. The improvement-to-regression ratio is 219:14; for every question the adapter regresses, it newly answers roughly 15.6 questions correctly that the base did not.

\paragraph{Per-tier and per-regulator breakdown.} Tier~2 numeric questions ($n{=}386$) gain $+24.6$~pp; Tier~3 heading-extractive questions ($n{=}278$) gain $+39.5$~pp. The adapter dominates on the heading-extractive tier where verbose paragraph quotation is the right deliverable. By regulator, RBI questions gain $+31.5$~pp; SEBI questions gain $+30.1$~pp---both regulators benefit, with effectively equal magnitude.

\subsection{BFSI recall paired ($n{=}214$)}
\label{sec:results-recall}

\begin{table}[h]
  \caption{No-context recall paired evaluation, $n{=}214$. Substring is uninformative on this task (near-zero for all modes); token F1 is the primary metric. The recipe transfers to a structurally different task type at $p{=}1.50\times10^{-16}$.}
  \label{tab:recall}
  \centering
  \begin{tabular}{lcc}
    \toprule
    Mode & F1 mean & Wilcoxon vs base \\
    \midrule
    Base & 0.158 & --- \\
    + \texttt{bfsi\_recall} & 0.219 ($+38.4\%$ rel.) & $p{=}1.50\times10^{-16}$ \\
    \bottomrule
  \end{tabular}
\end{table}

In 74.3\% of paired questions, adapter F1 strictly exceeds base F1. The Wilcoxon $p{=}1.50\times10^{-16}$ confirms that the recipe generalises to a structurally different task type (no-context recall, attention-only LoRA, shorter sequence length).

\subsection{IndiaFinBench OOD ($n{=}324$)}

\begin{table}[h]
  \caption{IndiaFinBench overall performance for \texttt{bfsi\_extract}; Gemini 2.5 Flash zero-shot value from \citet{pall2026indiafinbench}. The 57.6~pp gap is the predicted out-of-distribution failure mode of a fine-tuned adapter under question-style shift.}
  \label{tab:indiafin-overall}
  \centering
  \begin{tabular}{lccc}
    \toprule
    Metric & Score & Wilson 95\% CI & vs Gemini Flash \\
    \midrule
    Substring & \textbf{32.1\%} & [27.3, 37.4] & 89.7\% \\
    Normalised match & 32.7\% & [27.8, 38.0] & --- \\
    Token F1 (mean) & 0.288 & --- & --- \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \caption{IndiaFinBench per task type. Contradiction-detection substring scores are inflated by yes/no lexical overlap and contradicted by a near-zero F1 of 0.015; the model is not actually performing the task.}
  \label{tab:indiafin-tasks}
  \centering
  \begin{tabular}{lcccc}
    \toprule
    Task type & $n$ & Synapta sub.\ \% & Synapta F1 & Gemini Flash \\
    \midrule
    regulatory\_interpretation & 139 & 34.5\% & 0.347 & 93.1\% \\
    temporal\_reasoning        &  62 & 16.1\% & 0.366 & 88.5\% \\
    numerical\_reasoning       &  73 &  9.6\% & 0.297 & 84.8\% \\
    contradiction\_detection   &  50 & 78.0\% & 0.015 & 88.7\% \\
    \midrule
    Overall                    & 324 & 32.1\% & 0.288 & 89.7\% \\
    \bottomrule
  \end{tabular}
\end{table}

By regulator, the adapter scores SEBI 33.0\% ($n{=}273$) and RBI 27.5\% ($n{=}51$). The 57.6~pp gap between our 32.1\% on IndiaFinBench and Gemini Flash's reported 89.7\% is the predicted out-of-distribution failure mode of a fine-tuned adapter under question-style shift~\citep{liang2022helm}. Our 89.6\% on the in-distribution held-out set and Gemini's 89.7\% on IndiaFinBench are not directly comparable: different benchmarks, different scoring, different question styles. The apples-to-apples number for our adapter on IndiaFinBench is 32.1\%, and we report it without softening.

\subsection{Indian BFSI Benchmark v1 ($n{=}60$)}

\begin{table}[h]
  \caption{Released 60-question Indian BFSI Benchmark v1, mixed scoring. Paired McNemar adapter vs base $p{=}0.0313$. On the 30 substring-scored questions the adapter goes from 80\% to 100\%; on the 30 questions under the strict F1$\geq0.5$ cutoff both modes score 0\% (cutoff artefact discussed in text).}
  \label{tab:benchmark}
  \centering
  \begin{tabular}{lcccc}
    \toprule
    Mode & Score & Wilson 95\% CI & Substring & F1 \\
    \midrule
    Base & 40.0\% (24/60) & [28.6, 52.6] & 76.7\% & 0.122 \\
    + \texttt{bfsi\_extract} & \textbf{50.0\%} (30/60) & [37.7, 62.3] & 98.3\% & 0.157 \\
    \bottomrule
  \end{tabular}
\end{table}

On the 30 substring-scored questions, the adapter goes from 80\% to a perfect 100\%, a clean +20~pp lift. On the 30 questions scored under the strict F1$\geq0.5$ cutoff, both base and adapter score 0\%, despite mean F1 rising from $\sim0.12$ to $\sim0.16$. This is a metric artefact: a 1--2 sentence gold span almost never reaches F1$\geq0.5$ against a full-paragraph quotation, and the adapter is configured to quote the verbatim regulator paragraph---a citation-faithful deliverable preferred for compliance audits. The artefact motivates a v2 scoring revision (e.g., F1$\geq0.3$ or sentence-overlap variant for Tier~3) which we leave to a follow-up release.

\section{Limitations and negative findings}

We summarise key limitations; these scope our claims and are intended to prevent over-generalisation.

\begin{itemize}
  \item \textbf{Regulator coverage.} Our corpus covers RBI and SEBI only. India's BFSI regulatory perimeter additionally includes IRDAI (insurance), PFRDA (pensions), and the FEMA framework (foreign exchange). We do not claim multi-regulator coverage.
  \item \textbf{Small-$n$ on the released benchmark.} Benchmark v1 has $n{=}60$. The McNemar test on adapter vs base is $p{=}0.0313$---positive but marginal at $\alpha{=}0.05$. We treat the $n{=}664$ document-disjoint paired evaluation as the load-bearing statistical claim and the released benchmark as an open artefact for community extension.
  \item \textbf{No frontier-API comparison at scale.} A frontier-API paired comparison at $n{=}60$ against Anthropic Claude, OpenAI GPT-4o, and Google Gemini is left to v1.1 of the benchmark; we deliberately do not report any frontier comparison in this paper.
  \item \textbf{Benchmark-incomparability caveat.} We deliberately do not report ``our adapter 89.6\% $\approx$ Gemini Flash 89.7\%''. These are different benchmarks; the apples-to-apples numbers are 89.6\% vs base 58.7\% on our held-out, and 32.1\% vs Gemini's 89.7\% on IndiaFinBench.
  \item \textbf{F1 cutoff metric artefact.} The F1$\geq0.5$ cutoff in Benchmark v1 is too strict for verbose paragraph-extraction style; a revised scoring would likely yield non-zero scores. We document the artefact in v1 and plan a revised scorer for v2.
  \item \textbf{Single base model.} All BFSI results are on Nemotron-Nano-30B-A3B 4-bit. Multi-base replication (e.g., Qwen3-32B, Llama-3-70B 4-bit) is not reported here.
  \item \textbf{Long-context behaviour at the OOD boundary.} IndiaFinBench sample contexts (up to 8k tokens) exceed our training context (1024). We did not retrain at longer context; long-context degradation may inflate the OOD gap by an unknown amount.
  \item \textbf{Document-disjoint vs random-row ablation not run.} We argue the protocol on first-principles grounds; an ablation showing the inflation a random-row split would have produced on the same 4{,}477 pairs is left for a follow-up.
\end{itemize}

\section{Discussion and future work}
\label{sec:discussion}

\subsection{Reframing the OOD gap}

The natural reading of a $-57.6$~pp gap to Gemini Flash on IndiaFinBench is ``your model is much worse.'' We argue the more accurate reading is: a model fine-tuned on one customer's regulatory corpus does not transfer to a different question style on the same domain---the failure mode predicted by OOD evaluation work such as HELM~\citep{liang2022helm}. This is also the structural argument for per-customer adapters: if a single fine-tuned model could handle every Indian-BFSI customer's questions, then a single frontier API call could too, and the sovereignty problem would be vacuous. The corollary is operational: each customer should receive their own adapter, trained on their own document corpus via the deterministic pipeline reported here, with paired held-out evaluation against the base.

\subsection{Evaluation discipline as the contribution}

We position this paper as an evaluations-and-datasets contribution rather than a methods paper. The pipeline uses well-known primitives (LoRA, QLoRA, regex extraction, McNemar, Wilcoxon). The contribution is in their disciplined assembly: deterministic QA construction without LLM paraphrase, document-disjoint splitting before any pair is constructed, paired evaluation against the same base in three independent settings, and honest disclosure of the OOD result rather than an inflated cross-benchmark headline. The released artefacts---corpus manifests, 10-check validator, document-disjoint split manifest, 60-question benchmark with gated scorer, and per-question prediction JSONLs---are intended to make the protocol reusable for other Indian-BFSI customer corpora and, with minor scraper modifications, for other sovereign-regulatory NLP settings (e.g., other South Asian and Southeast Asian regulators with similar Master Direction / Master Circular document structures).

\subsection{Future work}

Concrete near-term extensions include:

\begin{enumerate}
  \item \textbf{Multi-regulator extension.} Extending the pipeline to IRDAI, PFRDA, and FEMA is likely to yield a larger train corpus and shrink the IndiaFinBench OOD gap on regulatory-interpretation tasks.
  \item \textbf{Frontier comparison at $n{=}60$ (v1.1).} Running Anthropic Claude, OpenAI GPT-4o, and Google Gemini against Benchmark v1 with paired McNemar would establish a properly statistical frontier comparison.
  \item \textbf{Document-disjoint vs random-row ablation.} Re-splitting the same 4{,}477 raw pairs at the row level and re-evaluating would quantify the contamination inflation that document-disjoint splitting prevents.
  \item \textbf{Per-customer pilot.} Applying the pipeline to a real customer's internal compliance corpus and reporting the held-out paired McNemar from that customer's adapter is the obvious deployment validation.
  \item \textbf{Long-context retraining.} Extending \texttt{bfsi\_extract}'s context window to 8k tokens and re-running IndiaFinBench would disentangle style-shift from length-shift in the OOD gap.
  \item \textbf{Multi-base replication.} Repeating \texttt{bfsi\_extract} on additional 30B-class bases (e.g., Qwen3-32B, Llama-3-70B 4-bit) to test base-invariance.
  \item \textbf{Benchmark v2.} A larger ($\sim200$-question) hand-curated benchmark with revised F1 cutoff (F1$\geq0.3$ or sentence-overlap for Tier~3) and additional regulators.
\end{enumerate}

\begin{ack}
We thank the maintainers of the RBI and SEBI public document portals for keeping regulatory text openly available; the IndiaFinBench authors for releasing a contemporary public benchmark with expert annotations; and the open-source authors of LoRA, QLoRA, PEFT, and \texttt{bitsandbytes}. This work received no external funding. The authors declare no competing interests.
\end{ack}

{
\small

\bibliographystyle{plainnat}

\begin{thebibliography}{9}

\bibitem[Dettmers et al.(2023)]{dettmers2023qlora}
T.~Dettmers, A.~Pagnoni, A.~Holtzman, and L.~Zettlemoyer.
\newblock {QLoRA}: Efficient finetuning of quantized {LLMs}.
\newblock In \emph{Advances in Neural Information Processing Systems (NeurIPS)}, 2023.

\bibitem[Hu et al.(2021)]{hu2021lora}
E.~J. Hu, Y.~Shen, P.~Wallis, Z.~Allen-Zhu, Y.~Li, L.~Wang, and W.~Chen.
\newblock {LoRA}: Low-rank adaptation of large language models.
\newblock \emph{arXiv preprint arXiv:2106.09685}, 2021.

\bibitem[Liang et al.(2022)]{liang2022helm}
P.~Liang, R.~Bommasani, T.~Lee, D.~Tsipras, D.~Soylu, M.~Yasunaga, et~al.
\newblock Holistic evaluation of language models.
\newblock \emph{arXiv preprint arXiv:2211.09110}, 2022.

\bibitem[Magar and Schwartz(2022)]{magar2022data}
I.~Magar and R.~Schwartz.
\newblock Data contamination: From memorization to exploitation.
\newblock In \emph{Proceedings of the 60th Annual Meeting of the Association for Computational Linguistics (ACL)}, 2022.

\bibitem[Pall et al.(2026)]{pall2026indiafinbench}
S.~Pall et~al.
\newblock IndiaFinBench: An evaluation benchmark for large language model performance on Indian financial regulatory text.
\newblock \emph{arXiv preprint arXiv:2604.19298}, 2026.

\end{thebibliography}
}

\appendix

\section{Reproducibility artefacts}

\paragraph{Released artefacts.} The corpus is reconstructable from the published RBI and SEBI manifests; the deterministic 3-tier QA pipeline, the 10-check validator, and the document-disjoint split manifest (\texttt{data/rbi\_corpus/qa/split\_manifest\_v2.json}) are released with the paper. The Indian BFSI Benchmark v1 ships with \texttt{questions.jsonl}, \texttt{scoring.py} (with built-in Wilson confidence intervals and paired McNemar), \texttt{README.md}, \texttt{LICENSE.md} (CC-BY-SA-4.0), and \texttt{dataset-metadata.json}. Per-question prediction JSONLs are included for the $n{=}664$ extract, $n{=}214$ recall, $n{=}324$ IndiaFinBench, and $n{=}60$ Benchmark~v1 evaluations, alongside their corresponding McNemar contingency tables and Wilcoxon outputs.

\paragraph{Decoding configuration.} All paired evaluations share temperature $0$, identical \texttt{max\_new\_tokens} budgets, identical prompt templates, and identical seeds. We do not re-tune decoding for any single benchmark.

\paragraph{Hardware.} All training and evaluation reported in this paper was run on a single consumer RTX~5090 with 32~GB VRAM. Adapter training is 3h~28min wall-clock at 17.81~GB peak VRAM; the IndiaFinBench evaluation is 73~min.

\paragraph{Broader impacts.} Positive impacts include lowering the cost and friction of regulator-faithful citation lookup for Indian-BFSI compliance teams, with on-premise deployment compatible with data-localisation constraints. Negative impacts include the risk that a fine-tuned adapter is deployed without paired evaluation in a setting whose question style differs from training---our IndiaFinBench result quantifies that failure mode and motivates the paired-evaluation protocol we recommend before any deployment. The released benchmark and scorer are intended to make such paired evaluation cheap to run.

\newpage
\input{checklist.tex}

\end{document}
